Beam-Helicity and Beam-Charge Asymmetries \\ in the Bethe-Heitler and DVCS processes on \\ an Unpolarised Proton Target \underline{Paper Tag: PDVCS0607 (v1.3)} \underline{Drafting Committee: DC90 (\today)}

### 2011-09-07 summary.tex
Small changes

Beam-helicity and beam-charge asymmetries in the azimuthal distribution of real photons from leptoproduction on an unpolarised hydrogen target have been presented. These asymmetries were extracted from a new unpolarised hydrogen data set taken during the 2006 and 2007 operating periods of H{\sc ermes}. Analogous asymmetry amplitudes were extracted previously from hydrogen data obtained during the 1996-2005 experimental period as described in Ref.~\cite{Air09}. A comparison of the amplitudes extracted from these independent data sets has shown that they are compatible and the asymmetry amplitudes can therefore be extracted from the complete 1996-2007 event sample. \red{The asymmetry amplitudes extracted from the combined data set are the statistically most precise DVCS measurements presented by H{\sc ermes}. There is} a strong signal in the first harmonic of the interference contribution to the beam-helicity asymmetry. There are non-zero amplitudes in the zeroth and first harmonics of the beam-charge asymmetry. Each of the higher order asymmetry amplitudes is consistent with zero. The combined results are compared to calculations from ongoing work to fit GPD models to experimental data. All data amplitudes are also presented as projections in $-t$ in bins of $x_{\textrm{B}}$. No additional features are observed in any particular $x_{\textrm{B}}$-bin.

>>>>

Beam-helicity and beam-charge asymmetries in the azimuthal distribution of real photons from leptoproduction on an unpolarised hydrogen target have been presented. These asymmetries were extracted from a new unpolarised hydrogen data set taken during the 2006 and 2007 operating periods of H{\sc ermes}. Analogous asymmetry amplitudes were extracted previously from hydrogen data obtained during the 1996-2005 experimental period as described in ref.~\cite{Air09}. A comparison of the amplitudes extracted from these independent data sets has shown that they are compatible and the asymmetry amplitudes can therefore be extracted from the complete 1996-2007 event sample. \red{The asymmetry amplitudes extracted from the combined data set are the statistically most precise DVCS measurements presented by H{\sc ermes}. There is} a strong signal in the first harmonic of the interference contribution to the beam-helicity asymmetry. There are non-zero amplitudes in the zeroth and first harmonics of the beam-charge asymmetry. Each of the higher order asymmetry amplitudes is consistent with zero. The combined results are compared to calculations from ongoing work to fit GPD models to experimental data. All data amplitudes are also presented as projections in $-t$ in bins of $x_{\textrm{B}}$. No additional features are observed in any particular $x_{\textrm{B}}$-bin.

### 2011-09-07 theory.tex
Small changes

\section{Introduction}
Generalised Parton Distributions (GPDs)
\cite{Mue94,Ji97a,Rad97} encompass the familiar parton
distribution functions (PDFs) and nucleon form factors to provide a
comprehensive description of the structure of the nucleon.
Although GPDs are difficult to access experimentally, the H{\sc ermes}
collaboration has previously published results
\cite{Air01,Air06,Air08,Air09,Air10, Air10a, Air10b, Air11} on
the leptoproduction of real photons from a nucleon or nucleus that can
be used to constrain GPDs via Compton Form Factors (CFFs). A thorough
description of the nucleon in terms of GPDs would allow the deduction
of the total angular momentum of partons in the nucleon, and the
construction of a longitudinal-momentum-dissected transverse spatial map of parton positions.

Generalised parton distributions depend upon four kinematic variables: the
Mandelstam variable $t=(p-p^{\prime})^2$, which is the squared momentum
transfer to the target nucleon in the scattering process with $p$ ($p^{\prime}$)
representing the initial (final) four-momentum of the proton; the average
fraction $x$ of the nucleon's longitudinal momentum carried by the active
quark throughout the scattering process; half the difference of
  the fractions of the nucleon's longitudinal momentum carried
by the active quark at the start and end of the process, written as
the skewness $\xi$; and the square $-Q^2=q^2$ of the four-momentum of
the virtual photon that mediates the lepton-proton scattering
process. In the Bjorken limit of $Q^2\rightarrow\infty$ with fixed
$t$, the skewness $\xi$ is related to the Bjorken variable
$x_{\textrm{B}}=\frac{-q^2}{2p\cdot q}$ as
$\xi\approx\frac{x_\textrm{B}}{2-x_\textrm{B}}$. The results are presented
as a function of $x_{\textrm{B}}$ because there is no consensus on an experimentally observable representation of $\xi$. 
\begin{figure}
\begin{center}
\subfigure[DVCS]{\includegraphics[width=0.45\textwidth]{dvcs_both}}
\subfigure[Bethe Heitler]{\includegraphics[width=0.45\textwidth]{bh_both}}
\caption[DVCS and Bethe Heitler hand bag diagram.]{(a): The DVCS process in
which an electron/positron ($e$) interacts with a quark in the nucleon
($p$) via a virtual photon ($\gamma^\ast$). The quark is found in the
nucleon with longitudinal momentum fraction $x+\xi$ and emits a real
photon ($\gamma$). The quark is absorbed by the nucleon with
longitudinal momentum fraction
$x-\xi$. (b): The Bethe-Heitler process, i.e. the emission of a real photon from the scattering or
scattered lepton. It has the same initial and final states as DVCS.}
\label{spin}
\end{center}
\end{figure}
Exclusive leptoproduction of real photons
($e\,p\,\rightarrow\,e'\,p'\,\gamma$) arises from
two experimentally indistinguishable processes shown in
Fig.~\ref{spin}: the Deeply Virtual Compton Scattering (DVCS) process,
which is the emission of a real photon by the struck quark from the nucleon, and the Bethe-Heitler (BH) process, which is elastic lepton-proton scattering with the
emission of a Bremsstrahlung photon by the lepton. 
The BH process is calculable in the QED framework; this process is
dominant at the kinematic conditions of the H{\sc ermes} experiment, but the
scattering amplitudes of the two processes interfere and the large BH amplitude
amplifies the contribution of the DVCS amplitude to the interference term. 
It is through the study of this interference term at H{\sc ermes} that
useful information for the constraint of certain GPDs can be obtained \cite{Bel02b}.

>>>>

\section{Introduction}
Generalised Parton Distributions (GPDs)
\cite{Mue94,Ji97a,Rad97} encompass the familiar Parton
Distribution Functions (PDFs) and nucleon Form Factors (FFs) to provide a
comprehensive description of the structure of the nucleon.
Although GPDs are difficult to access experimentally, the H{\sc ermes}
collaboration has previously published results
\cite{Air01,Air06,Air08,Air09,Air10, Air10a, Air10b, Air11} on
the leptoproduction of real photons from a nucleon or nucleus that can
be used to constrain GPDs via Compton Form Factors (CFFs). A thorough
description of the nucleon in terms of GPDs would allow the deduction
of the total angular momentum of partons in the nucleon, and the
construction of a longitudinal-momentum-dissected transverse spatial map of parton positions.

Generalised parton distributions depend upon four kinematic variables: the
Mandelstam variable $t=(p-p^{\prime})^2$, which is the squared momentum
transfer to the target nucleon in the scattering process with $p$ ($p^{\prime}$)
representing the initial (final) four-momentum of the proton; the average
fraction $x$ of the nucleon's longitudinal momentum carried by the active
quark throughout the scattering process; half the difference of
  the fractions of the nucleon's longitudinal momentum carried
by the active quark at the start and end of the process, written as
the skewness $\xi$; and the square $-Q^2=q^2$ of the four-momentum of
the virtual photon that mediates the lepton-proton scattering
process. In the Bjorken limit of $Q^2\rightarrow\infty$ with fixed
$t$, the skewness $\xi$ is related to the Bjorken variable
$x_{\textrm{B}}=\frac{-q^2}{2p\cdot q}$ as
$\xi\approx\frac{x_\textrm{B}}{2-x_\textrm{B}}$. The results are presented
as a function of $x_{\textrm{B}}$ because there is no consensus on an experimentally observable representation of $\xi$. 
\begin{figure}
\begin{center}
\subfigure[DVCS]{\includegraphics[width=0.45\textwidth]{dvcs_both}}
\subfigure[Bethe Heitler]{\includegraphics[width=0.45\textwidth]{bh_both}}
\caption[DVCS and Bethe Heitler hand bag diagram.]{(a): The DVCS process in
which an electron/positron ($e$) interacts with a quark in the nucleon
($p$) via a virtual photon ($\gamma^\ast$). The quark is found in the
nucleon with longitudinal momentum fraction $x+\xi$ and emits a real
photon ($\gamma$). The quark is absorbed by the nucleon with
longitudinal momentum fraction
$x-\xi$. (b): The Bethe-Heitler process, i.e. the emission of a real photon from the scattering or
scattered lepton. It has the same initial and final states as DVCS.}
\label{spin}
\end{center}
\end{figure}
Exclusive leptoproduction of real photons
($e\,p\,\rightarrow\,e'\,p'\,\gamma$) arises from
two experimentally indistinguishable processes shown in
figure~\ref{spin}: the Deeply Virtual Compton Scattering (DVCS) process,
which is the emission of a real photon by the struck quark from the nucleon, and the Bethe-Heitler (BH) process, which is elastic lepton-proton scattering with the
emission of a Bremsstrahlung photon by the lepton. 
The BH process is calculable in the QED framework; this process is
dominant at the kinematic conditions of the H{\sc ermes} experiment, but the
scattering amplitudes of the two processes interfere and the large BH amplitude
amplifies the contribution of the DVCS amplitude to the interference term. 
It is through the study of this interference term at H{\sc ermes} that
useful information for the constraint of certain GPDs can be obtained \cite{Bel02b}.
